Naturalness in U(1) extensions of the MSSM

### 2015-01-20 paper/e6ssm-tuning-paper.tex
explained the benchmark points as best I can for now.

To further illustrate point we looked at how the tuning compares for
low tuning benchmarks in the MSSM and E6SSM when the highscale
boundary condition, $M_X$ is raised.

In the MSSM the panel on the bottom right shows that the largest
tuning contributions come from $\Delta_{A_t}$ and $\Delta_{m_{H_u}^2}$
with the former being the larger sensitivity until $M_X\approx 10^8$
at which point $\Delta_{m_{H_u}^2}$ takes over.

The results of this are shown in
Fig.~\ref{Fig:BMs-varyMX} where in the top left panel we see that the
dependence of the E6SSM tuning on the cutoff is rather complicated
leading, such that any statement about which model is more tuned
depends on the cutoff scale considered.

E$_6$SSM BM1

The complicated shape of the E6SSM fine tuning curve can be understood
by looking at the indivual sensitivities shown in bottom left panel.
In this case one can see that the tuning is initially dominated by the
$\Delta_\lambda$ sensitivity, which contains some terms proprtional to
$M_Z^{\prime \, 2}$.  As the cutoff scale is raised however fine
tuning contributions from the stops and gluino start to feed into the
sensitivities, in particular $\Delta_{m_{H_u}^2}$ and $\Delta_\lambda$.
In the latter these contribution have the opposite sign to those from
$M_Z^\prime$ leading to this strange behaviour where $\Delta_\lambda$
passes through zero.  However in $\Delta_{m_{H_u}^2}$ dominating at
larger values of the cutoff scale we see the effect of the more common
fine tuning problem where the stops add large contributions to
$m_{H_u}^2$ , in this case leading to a large cancellation between
$m_{H_u}^2(M_X)$ and $m_{Q_3}^2(M_X)$ inside $m_{H_u}^2(M_{SUSY})$.

In this particular case the fine tuning of the E$_6$SSM benchmark becomes larger than that od the MSSM benchmark when $M_X\approx 10^{10}-10^{11}$ GeV. This is diue to ? large $A_t$ mixing?, $M_3$? different RGEs?

\begin{figure}
\begin{center}
\resizebox{!}{5.5cm}{%
\includegraphics{e6ssm_mssm_cutoff_scan_3_points_tuning_plot.png}} 
\resizebox{!}{5.5cm}{%
\includegraphics{mssm_cutoff_scan_tb10_fixed_at_ms_all_tuning_data.png}} 
\\
\resizebox{!}{5.5cm}{%
\includegraphics{e6ssm_light_zprime_cutoff_scan_tb10_fixed_at_ms_all_tuning_data.png}}
\resizebox{!}{5.5cm}{%
\includegraphics{e6ssm_light_zprime_cutoff_scan_tb10_fixed_at_ms_mssm_mixing_small_M3_all_tuning_data.png}}%%
\caption{{\bf Top left panel:} Scatter plot of fine tuning vs high scale $M_X$ for the three benchmark points given in (?).  {\bf Top right panel:}  Individual sensitivites of MSSM BM plotted against the high scale $M_X$ which give an overall tuning shown by the green line in top left panel.  {\bf Bottom left Panel:} Individual sensitivites of E6SSM BM1 plotted against the high scale $M_X$ which give an overall tuning shown by the blue line in top left panel.  {\bf Bottom right Panel:}Individual sensitivites of E6SSM BM2 plotted against the high scale $M_X$ which give an overall tuning shown by the red line in top left panel.    }
\label{Fig:BMs-varyMX}
\end{center}
\end{figure}

>>>>

To further illustrate this point we looked at how the tuning compares for
low tuning benchmarks in the MSSM and E6SSM when the highscale
boundary condition, $M_X$ is raised.  The results of this are shown in
Fig.~\ref{Fig:BMs-varyMX}.

In the top left panel one can see that the MSSM tuning (green)
steadily climbs as the cutoff scale is increased, as one would expect
when the tuning originates from large soft masses entering from the
RGEs. The panel on the top right confirms this, showing that the
largest tuning contributions come from $\Delta_{A_t}$ and
$\Delta_{m_{H_u}^2}$ with the former being the larger sensitivity
until $M_X\approx 10^8$ GeV at which point $\Delta_{m_{H_u}^2}$ takes
over, leading to the small kink in overall tuning which can be seen in
the green curve in the top left panel. In thus case we have chosen a
point with large mixing which reduce the MSSM tuning somewhat.  this
does not eliminate the tuning as one there is still a string
sensitivity to $A_t$, but we did find that increasing $A_t$ lead less
fine tuning overall.

The E$_6$SSM tunings are rather more complicated.  By comparing them
to the MSSM tuning one can see that the scale at which the parameters
are defined is very important for determining which point is more fine
tuned.  In the bottom left panel one can see that the dependence of
the E$_6$SSM tuning on the cutoff is rather complicated, illustrating
that any statement about which model is more tuned depends on the
cutoff scale considered.

For E$_6$SSM BM1 the fine tuning is shown by the blue curve in top
left panel of Fig.~\ref{Fig:BMs-varyMX} and the individual sensitivities
are given in the bottom left panel. The tuning actually reduces
initially as the cutoff is is increased from $20$ TeV.  This occurs
because the tuning is initially dominated by $\Delta_\lambda$ (shown in
solid blue in the bottom left panel) that contains some terms
proportional to $M_Z^{\prime \, 2}$, which currently dominate the
tuning.  However as the cutoff scale is increased contributions from
the soft masses become more important and these actually start to
cancel the large contribution to $\Delta_\lambda$ coming from
$M_{Z^\prime}$ until $\Delta_\lambda$ passes through zero.  However at
the same time these large soft masses also cause other sensitivities
to grow, leading the overall tuning to start rising with $M_X$
once. Nonetheless due to the lower stop masses and lower mixing [I'm
  guessing slightly: Dylan I need to be able to refer to numbers
  here!]  compared to both of the other points until $M_X \approx
10^{11}$ GeV.  After this the tuning from $M_3$ starts to dominate in
this point.  Although at $M_{SUSY}$ this point as the same $M_3$ value
as the MSSM point, in the E$_6$SSM $M_3$ will be larger at $M_X$ due
to the altered RGE running from exotic matter\footnote{This altered
  RGE running is a result of the exotic matter introduced to keep the
  $U(1)$ anomaly free rather than being a direct consequence of the
  extra $U(1)$. }.  At the same time larger $M_3$ at higher scales
reduces the soft squark masses at $M_X$.  Therefore the stop mass
contributions are ameliorated, compared to the MSSM, both by allowing
lighter stops at $M_{SUSY}$ and by the modified RGE running.
Nonetheless the stops still have an impact, pushing up the sensitivity
to $m_{H_u}^2$\footnote{$m_{H_u}^2(M_{SUSY})$ receives a positive
  contribution to it's mass from $m_{H_u}^2(M_X)$ and a negative
  contribution from $m_{Q_3}^2$. If $m_{H_u}^2(M_{SUSY})$ is held
  fixed so that then as the cutoff increases the values of these
  softmasses at $M_X$ will be larger and there we be a bigger
  cancellation between them, making the sensitivity of $M_Z$ to both of
  these parameters larger.}  showing the most commonly discussed
aspect of the little hierarchy problem.

By contrast the tuning for E$_6$SSM BM2 is very different.  This point
was chosen as it had much lighter stops and a gluino mass that is just
above the experimental limit of $1.4$ TeV \cite{}.  At $20$ TeV it was
not amongst the lowest tuned points, since at that scale the tree
level tuning from $M_{Z^\prime}$ dominates.  The reduction in $M_3$
means that $\Delta_{M_3}$ is substantially lower and only becomes the
dominant tuning very close to the GUT scale, giving a tuning at
$10^{16}$ GeV of $\approx 900$, which is below that of the other two
benchmark points.  Since large mixing was also used to reduce allow
lighter stops, we see initially a large contribution from
$\Delta_{A_t}$ as we did in the MSSM benchmark.  [Last thing that needs to be explained - why do $\Delta_{M_3}$ and $\Delta_{m_{H_u}^2}$ start to decrease at $M_X$.  I'm guessing something funny to do with their values and also  $m_{Q}^2$ and $M_3$ at the highscales.  but i need to see the numbers to know.]

\begin{figure}
\begin{center}
\resizebox{!}{5.5cm}{%
\includegraphics{e6ssm_mssm_cutoff_scan_3_points_tuning_plot.png}} 
\resizebox{!}{5.5cm}{%
\includegraphics{mssm_cutoff_scan_tb10_fixed_at_ms_all_tuning_data.png}} 
\\
\resizebox{!}{5.5cm}{%
\includegraphics{e6ssm_light_zprime_cutoff_scan_tb10_fixed_at_ms_all_tuning_data.png}}
\resizebox{!}{5.5cm}{%
\includegraphics{e6ssm_light_zprime_cutoff_scan_tb10_fixed_at_ms_mssm_mixing_small_M3_all_tuning_data.png}}%%
\caption{{\bf Top left panel:} Scatter plot of fine tuning vs high scale $M_X$ for the three benchmark points given in (?).  {\bf Top right panel:}  Individual sensitivites of MSSM BM plotted against the high scale $M_X$ which give an overall tuning shown by the green line in top left panel.  {\bf Bottom left panel:} Individual sensitivites of E6SSM BM1 plotted against the high scale $M_X$ which give an overall tuning shown by the blue line in top left panel.  {\bf Bottom right panel:}Individual sensitivites of E6SSM BM2 plotted against the high scale $M_X$ which give an overall tuning shown by the red line in top left panel.    }
\label{Fig:BMs-varyMX}
\end{center}
\end{figure}
